Gradient Descent Method in Learning

### 2020-08-30 publications/princeton.tex
Add files via upload

\documentclass[pdf,slideColor,colorBG]{prosper}

@@

%\overlays{7}{%
\begin{slide}{Outline of the talk}
\begin{itemize}
    \item Batch vs Online learning
    \item Gradient Descent Method in both settings
    \item How to do regularizations
    \item Lower and Upper Bounds
    \item Exponential Rates for Plug-in Classifiers
    \item ``Random Connections'' with Random Projections
    \item Future Directions
\end{itemize}

@@

%
%\overlays{2}{%
\begin{slide}{Where we start: Penalizations}

@@

\medskip
%\fromSlide{2}{and $\H=\H_K$ a reproducing kernel Hilbert space (RKHS).}
and $\H=\H_K$ a reproducing kernel Hilbert space (RKHS).
\end{slide}
%}

%
%\overlays{3}{
\begin{slide}{Why RKHS}

\begin{itemize}
        \item $\H_K$ can be dense in continuous function space $\C(X)$
        \item the gradient takes a simple form in $\H_K$:
        $$\grad_f V = V'_f(y,f(x)) K_x $$
        \item super-fast convergence rate for plug-in classifiers...
\end{itemize}
\end{slide}
%}

@@

%
%\overlays{2}{
\begin{slide}{Our Theoretical Goal}
Convergence of $(\hat{f}_t)\in \H_K$ and $(f_t)\in \H_K$ to the regression function
\[ f_\rho (x) := \E[y|x] \in \L2 \]
and its rates when $f_\rho$ takes some sparse form.

@@

But, $\L2$ is too large a space to search, so we need \emph{regularizations}.

@@

\begin{slide}{Sparsity of Regression Function}
We are going to assume that the regression function is sparse/smooth w.r.t. the following \emph{basis}
\begin{itemize}
    \item roughly speaking, kernel principle components,
    \item or more precisely, the eigenfunctions of the \emph{covariance operator} of $\rho_\X$ on $\H_K$.
\end{itemize}
\end{slide}
%
%\overlays{3}{
\begin{slide}{Covariance operator}
\begin{itemize}
    \item Define an integral operator
\[
\begin{array}{rcl}
L_K : \L2 & \to & \H_K \\
       f & \mapsto & \int_X f(x') K(x',\cdot) d \rho_X
\end{array}
\]
    \item The \emph{covariance operator}, is the restriction $L_K|_{\H_K}:\H_K\to \H_K$, i.e. $\E_x[ \<\ ,K_x\>K_x]$

    \item $L_K:\L2\to \L2$ compact $\Rightarrow$ orthonormal eigen-system $(\la_i,\phi_i)_{i\in \N}$, $\phi_i \in \L2\cap\H_K$ bi-orthogonal and
\[ \sum \la_i \leq \sup_{x\in \X} K(x,x) =: \kappa <\infty \]
\end{itemize}
\end{slide}
%}

%
%\overlays{2}{
\begin{slide}{Sparsity Assumption}
Assume that
\[ f_\rho = L_K^r g, \ \ \ \ g\in \L2, r>0 \]
i.e. $f_\rho$ has at least \emph{power-law decay} coordinates w.r.t. the basis of eigenfunctions of $L_K:\L2\to \L2$:
\[ f_\rho = \sum_i \la_i^r g_i \phi_i,  \ \ \ \sum \la_i <\infty, \sum g_i^2 <\infty\]

\medskip
Note: if $K$ is a stochastic density kernel, $L_K^r$ is used to construct diffusion wavelets
by Coifman et al.
\end{slide}
%}

@@

%
%\overlays{3}{
\begin{slide}{Minimax Lower Rates}
[Caponnetto-DeVito'05] The minimax lower rate:
\[
\begin{array}{rl}
& \liminf_{t\to \infty} \inf_{\z_t \mapsto f_t } \sup_{\rho\in \P(b,r)} \\
& \Prob \left\{\z_t \in \Z^t: \frac{\|f_t - f_\rho\|_2}{t^{-\frac{rb}{2rb+1}}} >C \right\} = 1
\end{array}
\]
where the $\inf$ is taken over all algorithms mapping $(z_i)_1^t\mapsto f_t$.

@@

\begin{itemize}
    \item The $\rho$ in $\sup_{\rho\in \P(b,r)}$ depends on sample size $t$!
    \item Suitable for online learning, instead of batch learning.
\end{itemize}
\end{slide}
%}

%
%\overlays{2}{
\begin{slide}{Individual Lower Rates}
[Caponnetto-DeVito'05] The individual lower rate: for each $B>b$,
\[\inf_{((z_i)_1^t\mapsto f_t)_{t\in \N}} \sup_{\rho \in \P(b,r)} \limsup_{t\to \infty} \frac{\|f_t- f_\rho\|_2}{t^{-\frac{rB}{2rB+1}}} > 0. \]

@@

Note: taking $b=1$ and $B=1$, it suggests \emph{eigenvalue independent} minimax and individual lower rates:
$$t^{-\frac{r}{2r+1}} $$
\end{slide}
%}

@@

%
%\overlays{2}{
\begin{slide}{Upper Bounds for Online Learning}
Theorem (Tarr\`es-Yao'06). Assume that $f_\rho = L_K^r g$ ($r>0$). There exist $\la_t$ and $\eta_t$ such that

@@

\emph{Note}: the upper rates \emph{saturate} when $r\geq 1$ and $r\geq 3/2$!
\end{slide}
%}

%
%\overlays{2}{
\begin{slide}{Breaking Saturation}
It is expected that with $\la_t=0$ and suitable choices $\eta_t\to 0$ and $\sum_t \eta_t =\infty$, one has
\[ \|f_t - f_\rho \|_2 \leq O(t^{-\frac{r}{2r+1}}) \]
for \emph{all $r>0$}.

@@

\emph{A Positive Answer}: Ying et al. (2006) give results suggesting its truth.
\end{slide}
%}

@@

%
%\overlays{4}{
\begin{slide}{Random Projection Perspective}
Given $\x_T\in \X^T$, define a sampling operator on $\H_K$
\[
\begin{array}{rcl}
S_{\x_T} : \H_K & \to & l_2(\x_T) \\
f & \mapsto & (f(x_i))_1^T = (\<f,K_{x_i}\>_K)_1^T
\end{array}
\]
\begin{itemize}
    \item $S_{\x_T} f$ takes $T$ random measurements/projections of $f$.
    \item Adjoint operator $S_{\x_T}^\ast \y = \frac{1}{T}\sum_{i=1}^T y_i K_{x_i}$.
    \item $S_{\x_T} S_{\x_T}^\ast$ is the Gram matrix $(K(x_i,x_j))^{T\times T}$.
\end{itemize}
\end{slide}
%}

%
%\overlays{2}{
\begin{slide}{Compressed Sensing}
\begin{itemize}
\item $f$ is sparse w.r.t. certain basis/frames (unknown)
\item $S_{\x_T}$ takes some random measurements of $f$ such that the Uniform Uncertainty Principle holds, or
equivalently, for small enough $T_0$ and all $T\leq T_0$, $S_{\x_T} S^\ast_{\x_T}$ has a \emph{uniform lower bound} (depending on the sparsity of $f$)
on the smallest eigenvalue.
\end{itemize}
\medskip

However in Learning, since
\[ \E[S^\ast_{\x_T} S_{\x_T}] = L_K|_{\H_K} \]
where $L_K$ is a compact operator with eigenvalues convergent to $0$, NO lower bound!
\end{slide}
%}

%
%\overlays{2}{
\begin{slide}{Learning vs. Compressed Sensing}
To control the \emph{condition number} (or smallest eigenvalue) of the Gram matrix $S_{\x_T}S_{\x_T}^\ast$:

@@

%\fromSlide{2}{
Moreover, there is another kind of ``condition number'' in machine learning: \\
\begin{center}
\emph{margin}
\end{center}
%}
\end{slide}
%}

@@

%
%\overlays{2}{
\begin{slide}{Margin and Random Projections}
[Balcan-Blum-Vempala'05] If $f\in \H_K$ has margin $\ga$, then with i.i.d. examples of number
\[ t \geq  \frac{8}{\epsilon} \max\left \{ \frac{1}{\ga^2}, \ln \frac{1}{\delta} \right \} \]
there is a $f_t$ such that with confidence $1-\delta$, $f_t$ has margin $\ga/2$ with
error $\epsilon$.

@@

\begin{itemize}
    \item In fact, $f_t$ can be realized by the Orthogonal Projections on to
    $\span \{K_i:1\leq i \leq t\}$.
\end{itemize}
\end{slide}
%}

>>>>

\documentclass[pdf,autumn,slideColor,colorBG]{prosper}

@@

\overlays{7}{%
\begin{slide}{Outline of the talk}
\begin{itemstep}
    \item Batch vs Online learning
    \item Gradient Descent Method in both settings
    \item How to do regularizations
    \item Lower and Upper Bounds
    \item Exponential Rates for Plug-in Classifiers
    \item ``Random Connections'' with Random Projections
    \item Future Directions
\end{itemstep}

@@

%
\overlays{2}{%
\begin{slide}{Where we start: Penalizations}

@@

\medskip
\fromSlide{2}{and $\H=\H_K$ a reproducing kernel Hilbert space (RKHS).}

%
\overlays{3}{
\begin{slide}{Why RKHS}

\begin{itemize}
        \item \fromSlide{1}{$\H_K$ can be dense in continuous function space $\C(X)$}
        \item \fromSlide{2}{the gradient takes a simple form in $\H_K$:
        $$\grad_f V = V'_f(y,f(x)) K_x $$}
        \item \fromSlide{3}{super-fast convergence rate for plug-in classifiers...}
\end{itemize}
\end{slide}
}

@@

%
\overlays{2}{
\begin{slide}{Our Theoretical Goal}
Convergence of $(\hat{f}_t)\in \H_K$ and $(f_t)\in \H_K$ to the regression function
\[ f_\rho (x) := \E[y|x] \in \L2 \]
and its rates when $f_\rho$ takes some sparse form.

@@

\fromSlide{2}{But, $\L2$ is too large a space to search, so we need \emph{regularizations}.}

@@

\begin{slide}{Sparsity of Regression Function}
We are going to assume that the regression function is sparse/smooth w.r.t. the following \emph{basis}
\begin{itemize}
    \item roughly speaking, kernel principle components,
    \item or more precisely, the eigenfunctions of the \emph{covariance operator} of $\rho_\X$ on $\H_K$.
\end{itemize}
\end{slide}
%
\overlays{3}{
\begin{slide}{Covariance operator}
\begin{itemstep}
    \item Define an integral operator
\[
\begin{array}{rcl}
L_K : \L2 & \to & \H_K \\
       f & \mapsto & \int_X f(x') K(x',\cdot) d \rho_X
\end{array}
\]
    \item The \emph{covariance operator}, is the restriction $L_K|_{\H_K}:\H_K\to \H_K$, i.e. $\E_x[ \<\ ,K_x\>K_x]$

    \item $L_K:\L2\to \L2$ compact $\Rightarrow$ orthonormal eigen-system $(\la_i,\phi_i)_{i\in \N}$, $\phi_i \in \L2\cap\H_K$ bi-orthogonal and
\[ \sum \la_i \leq \sup_{x\in \X} K(x,x) =: \kappa <\infty \]
\end{itemstep}
\end{slide}
}

%
\overlays{2}{
\begin{slide}{Sparsity Assumption}
Assume that
\[ f_\rho = L_K^r g, \ \ \ \ g\in \L2, r>0 \]
i.e. $f_\rho$ has at least \emph{power-law decay} coordinates w.r.t. the basis of eigenfunctions of $L_K:\L2\to \L2$:
\[ f_\rho = \sum_i \la_i^r g_i \phi_i,  \ \ \ \sum \la_i <\infty, \sum g_i^2 <\infty\]

\medskip
\fromSlide{2}{Note: if $K$ is a stochastic density kernel, $L_K^r$ is used to construct diffusion wavelets
by Coifman et al. }
\end{slide}
}

@@

%
\overlays{3}{
\begin{slide}{Minimax Lower Rates}
[Caponnetto-DeVito'05] The minimax lower rate:
\[
\begin{array}{rl}
& \liminf_{t\to \infty} \inf_{\z_t \mapsto f_t } \sup_{\rho\in \P(b,r)} \\
& \Prob \left\{\z_t \in \Z^t: \frac{\|f_t - f_\rho\|_2}{t^{-\frac{rb}{2rb+1}}} >C \right\} = 1
\end{array}
\]
where the $\inf$ is taken over all algorithms mapping $(z_i)_1^t\mapsto f_t$.

@@

\begin{itemize}
    \fromSlide{2}{\item The $\rho$ in $\sup_{\rho\in \P(b,r)}$ depends on sample size $t$!}
    \fromSlide{3}{\item Suitable for online learning, instead of batch learning.}
\end{itemize}
\end{slide}
}

%
\overlays{2}{
\begin{slide}{Individual Lower Rates}
[Caponnetto-DeVito'05] The individual lower rate: for each $B>b$,
\[\inf_{((z_i)_1^t\mapsto f_t)_{t\in \N}} \sup_{\rho \in \P(b,r)} \limsup_{t\to \infty} \frac{\|f_t- f_\rho\|_2}{t^{-\frac{rB}{2rB+1}}} > 0. \]

@@

\fromSlide{2}{Note: taking $b=1$ and $B=1$, it suggests \emph{eigenvalue independent} minimax and individual lower rates:
$$t^{-\frac{r}{2r+1}} $$
}
\end{slide}
}

@@

%
\overlays{2}{
\begin{slide}{Upper Bounds for Online Learning}
Theorem (Tarr\`es-Yao'06). Assume that $f_\rho = L_K^r g$ ($r>0$). There exist $\la_t$ and $\eta_t$ such that

@@

\fromSlide{2}{\emph{Note}: the upper rates \emph{saturate} when $r\geq 1$ and $r\geq 3/2$!}
\end{slide}
}

%
\overlays{2}{
\begin{slide}{Breaking Saturation}
It is expected that with $\la_t=0$ and suitable choices $\eta_t\to 0$ and $\sum_t \eta_t =\infty$, one has
\[ \|f_t - f_\rho \|_2 \leq O(t^{-\frac{r}{2r+1}}) \]
for \emph{all $r>0$}.

@@

\fromSlide{2}{\emph{A Positive Answer}: Ying et al. (2006) give results suggesting its truth.}
\end{slide}
}

@@

%
\overlays{4}{
\begin{slide}{Random Projection Perspective}
Given $\x_T\in \X^T$, define a sampling operator on $\H_K$
\[
\begin{array}{rcl}
S_{\x_T} : \H_K & \to & l_2(\x_T) \\
f & \mapsto & (f(x_i))_1^T = (\<f,K_{x_i}\>_K)_1^T
\end{array}
\]
\begin{itemize}
    \fromSlide{2}{\item $S_{\x_T} f$ takes $T$ random measurements/projections of $f$.}
    \fromSlide{3}{\item Adjoint operator $S_{\x_T}^\ast \y = \frac{1}{T}\sum_{i=1}^T y_i K_{x_i}$.}
    \fromSlide{4}{\item $S_{\x_T} S_{\x_T}^\ast$ is the Gram matrix $(K(x_i,x_j))^{T\times T}$.}
\end{itemize}
\end{slide}
}

%
\overlays{2}{
\begin{slide}{Compressed Sensing}
\begin{itemize}
\item $f$ is sparse w.r.t. certain basis/frames (unknown)
\item $S_{\x_T}$ takes some random measurements of $f$ such that the Uniform Uncertainty Principle holds, or
equivalently, for small enough $T_0$ and all $T\leq T_0$, $S_{\x_T} S^\ast_{\x_T}$ has a \emph{uniform lower bound} (depending on the sparsity of $f$)
on the smallest eigenvalue.
\end{itemize}
\medskip

\fromSlide{2}{However in Learning, since
\[ \E[S^\ast_{\x_T} S_{\x_T}] = L_K|_{\H_K} \]
where $L_K$ is a compact operator with eigenvalues convergent to $0$, NO lower bound!}
\end{slide}
}

%
\overlays{2}{
\begin{slide}{Learning vs. Compressed Sensing}
To control the \emph{condition number} (or smallest eigenvalue) of the Gram matrix $S_{\x_T}S_{\x_T}^\ast$:

@@

\fromSlide{2}{
Moreover, there is another kind of ``condition number'' in machine learning: \\
\begin{center}
\emph{margin}
\end{center}
}
\end{slide}
}

@@

%
\overlays{2}{
\begin{slide}{Margin and Random Projections}
[Balcan-Blum-Vempala'05] If $f\in \H_K$ has margin $\ga$, then with i.i.d. examples of number
\[ t \geq  \frac{8}{\epsilon} \max\left \{ \frac{1}{\ga^2}, \ln \frac{1}{\delta} \right \} \]
there is a $f_t$ such that with confidence $1-\delta$, $f_t$ has margin $\ga/2$ with
error $\epsilon$.

@@

\begin{itemize}
    \fromSlide{2}{\item In fact, $f_t$ can be realized by the Orthogonal Projections on to
    $\span \{K_i:1\leq i \leq t\}$.}
\end{itemize}
\end{slide}
}
